\textbf{Unshuffling Permutations}

### 2016-04-06 slides/slides.tex
Restructuration

@@

\begin{frame}
  \frametitle{Words}

  \begin{definition}
    The shuffle operator on words is defined inductively as follows:
    \begin{align*}
      u \shuffle \epsilon &= \epsilon \shuffle u = \{u\}
      \\
      (ua \shuffle vb) &= (ua \shuffle v)b \cup (u \shuffle vb)a
    \end{align*}
  \end{definition}

  \begin{block}{Key results}
    \begin{itemize}
      \item
      Given two words $v_1$ and $v_2$, the set $v_1 \shuffle v_2$
      can be computed in
      $O\left((m_1 + m_2) \; \binom{m_1 + m_2}{m_1}\right)$
      time.

      \item
      Given three words $u$, $v_1$, and $v_2$, deciding if $u$ is a shuffle
      of $v_1$ and $v_2$ can be done in
      $O\left(n^2 / \log(n)\right)$ time.

      \item
      Given a word $u$, deciding if there is a word $v$ such that $u$ is
      in $v \shuffle v$ is \NPC.
    \end{itemize}
  \end{block}

In a general way, for any $z \in C$,
\begin{equation*}
    \Delta(x) = \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
\end{equation*}
\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Equivalence between products and coproducts}
If $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$ is a product with
structure coefficients $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$,
we can define its \alert{dual coproduct} $\Delta$ by setting
\begin{equation*}
    \Delta(x) := \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
\end{equation*}
\medskip

Conversely, if $\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$ is
a coproduct with structure coefficients $\lambda_{x, y}^z \in \QQ$,
$x, y, z \in C$, we can define its \alert{dual product} $\cdot$ by
setting
\begin{equation*}
    x \cdot y := \sum_{z \in C} \lambda_{x, y}^z \; z.
\end{equation*}
\medskip

\begin{block}{Idea}
    Knowing how to break a combinatorial object is equivalent
    to knowing how to combine these.
\end{block}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Shuffling permutations}
\begin{block}{Definition}
    The \alert{shuffling product of permutations} is the product
    $\SHUFFLE$ on $\QQ[S]$ defined as the dual product of the unshuffling
    product of permutations.
\end{block}

\begin{block}{Example}
\begin{equation*}\begin{split}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2 (1342) + 2 (1423) + 3 (1432) +
    2134 + 2 (2314) \\
    & + 3 (2341) + 2413 + 2 (2431) + 2 (3124) + 3142 +
    3 (3214) + 2 (3241) \\
    & + 3421 + 3 (4123) + 2 (4132) + 2 (4213) + 4231 + 4312
\end{split}\end{equation*}
\end{block}
\medskip

Product first introduced for the study of some Hopf algebras~\BIB{Vargas}{2014}.
\medskip

Similar to the convolution product and the shifted
shuffle product \BIB{Duchamp, Hivert, Thibon}{2002}.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Some properties of square permutations}
The first numbers of square permutations are, size by size,
\begin{equation*}
    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
\end{equation*}
\medskip

\begin{block}{Proposition}
    Let $\pi$ be a square permutation and $\sigma$ be a square root of
    $\pi$. Then,
    \begin{enumerate}
        \item $\widetilde{\pi}$ is a square and
        $\widetilde{\sigma}$ is one of its square roots;
        \item $\bar \pi$ is a square and $\bar \sigma$ is one of
        its square roots;
        \item $\pi^{-1}$ is a square and $\sigma^{-1}$ is one of
        its square roots.
    \end{enumerate}
\end{block}
\medskip

>>>>

\begin{frame}
  \frametitle{Key results}

Given two words $v_1$ and $v_2$, the set $v_1 \shuffle v_2$
can be computed in
\begin{equation*}
    O\left((|v_1| + |v_2|) \; \binom{|v_1| + |v_2|}{|v_1|}\right)
\end{equation*}
time \BIB{Spehner}{1986}.
\bigskip

Given three words $u$, $v_1$, and $v_2$, deciding if $u$ is a
shuffle of $v_1$ and $v_2$ can be done in
\begin{equation*}
O\left(\frac{|u|^2}{\log(|u|)}\right)
\end{equation*}
time \BIB{van Leeuwen, Nivat}{1982}.
\bigskip

Given a word $u$, deciding if there is a word $v$ such that $u$ is
in $v \shuffle v$ is \NPC{} \BIB{Rizzi, Vialette}{2013}
\BIB{Buss, Soltys}{2014}.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{???}
JE PENSE QU'IL FAUDRAIT PARLER ICI DES COUPLAGES PARFAITS POUR LES MOTS.
\medskip

ÇA PERMET DE FIXER QUELQUES DEFINITIONS POUR PREPARER LE TERRAIN POUR
NOTRE EXTENSION POUR LES PERMU DANS LA PARTIE ALGO.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%o%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Main motivations}
Let $S_n$ be the set of permutations of size $n$.
\medskip

There are several products on permutations, analogs of the shuffle
product of words. Among these, there are
\begin{itemize}
    \item the shifted shuffle product \BIB{Duchamp, Hivert, Thibon}{2002};
    \medskip

    \item the convolution product \BIB{Duchamp, Hivert, Thibon}{2002};
    \medskip

    \item the supershuffle \BIB{Vargas}{2014}.
\end{itemize}
\medskip

\begin{block}{Main question}
    How the previous key results about the shuffle of words translate
    for the supershuffle of permutations?
\end{block}
\end{frame}

In a general way, for any $z \in C$,
\begin{equation*}
    \Delta(z) = \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
\end{equation*}
\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Equivalence between products and coproducts}
\begin{block}{Key idea}
    Knowing how to break a combinatorial object is equivalent
    to knowing how to combine these.
\end{block}
\medskip

Indeed, if $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$ is a product with
structure coefficients $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$,
we can define its \alert{dual coproduct} $\Delta$ by setting
\begin{equation*}
    \Delta(z) := \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
\end{equation*}
\medskip

Conversely, if $\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$ is
a coproduct with structure coefficients $\lambda_{x, y}^z \in \QQ$,
$x, y, z \in C$, we can define its \alert{dual product} $\cdot$ by
setting
\begin{equation*}
    x \cdot y := \sum_{z \in C} \lambda_{x, y}^z \; z.
\end{equation*}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Shuffling permutations}
\begin{block}{Definition}
    The \alert{shuffling product of permutations} (supershuffle) is the
    product $\SHUFFLE$ on $\QQ[S]$ defined as the dual product of the
    unshuffling product of permutations.
\end{block}
\medskip

\begin{block}{Proposition}
The permutations appearing in $\sigma \SHUFFLE \nu$ are exactly the
one obtained by shuffling two words $u$ and $v$ respectively
order-isomorphic to $\sigma$ and $\nu$.
\end{block}
\medskip

\begin{block}{Example}
\begin{equation*}\begin{split}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2 (1342) + 2 (1423) + 3 (1432) +
    2134 + 2 (2314) \\[.5em]
    & + 3 (2341) + 2413 + 2 (2431) + 2 (3124) + 3142 +
    3 (3214) + 2 (3241) \\[.5em]
    & + 3421 + 3 (4123) + 2 (4132) + 2 (4213) + 4231 + 4312
\end{split}\end{equation*}
\end{block}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Some properties of square permutations}
The first numbers of square permutations are, size by size,
\begin{equation*}
    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
\end{equation*}
%\medskip

\begin{block}{Proposition}
    Let $\pi$ be a square permutation and $\sigma$ be a square root of
    $\pi$. Then,
    \begin{enumerate}
        \item $\widetilde{\pi}$ is a square and
        $\widetilde{\sigma}$ is one of its square roots;
        \smallskip

        \item $\bar \pi$ is a square and $\bar \sigma$ is one of
        its square roots;
        \smallskip

        \item $\pi^{-1}$ is a square and $\sigma^{-1}$ is one of
        its square roots.
    \end{enumerate}
\end{block}
\medskip

@@
